\documentclass{article}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{hyperref}
\title{Introduction to Smart Pointers in C++}
\date{}
\begin{document}
\maketitle

\section*{What are Smart Pointers?}
Smart pointers are class templates in \texttt{<memory>} that automatically manage the lifetime of dynamically allocated objects. Unlike raw pointers, they prevent memory leaks by automatically deallocating memory when it's no longer needed.

\section*{Types of Smart Pointers}

\subsection*{1. \texttt{std::unique\_ptr}}
A \texttt{unique\_ptr} owns an object exclusively. It cannot be copied, only moved.

\textbf{Example:}
\begin{lstlisting}[language=C++]
#include <memory>
#include <iostream>

struct Foo {
    Foo() { std::cout << "Foo created\n"; }
    ~Foo() { std::cout << "Foo destroyed\n"; }
};

int main() {
    std::unique_ptr<Foo> p = std::make_unique<Foo>();
} // Foo is destroyed automatically here
\end{lstlisting}

\subsection*{2. \texttt{std::shared\_ptr}}
A \texttt{shared\_ptr} allows multiple pointers to share ownership of the same object. The object is deleted when the last \texttt{shared\_ptr} is destroyed.

\textbf{Example:}
\begin{lstlisting}[language=C++]
#include <memory>
#include <iostream>

int main() {
    auto p1 = std::make_shared<int>(42);
    auto p2 = p1; // reference count increases
    std::cout << *p2 << std::endl; // prints 42
} // object deleted when last shared_ptr goes out of scope
\end{lstlisting}

\subsection*{3. \texttt{std::weak\_ptr}}
A \texttt{weak\_ptr} is a non-owning reference to a \texttt{shared\_ptr} managed object. It is used to avoid cyclic dependencies.

\textbf{Example:}
\begin{lstlisting}[language=C++]
#include <memory>
#include <iostream>

int main() {
    std::shared_ptr<int> sp = std::make_shared<int>(100);
    std::weak_ptr<int> wp = sp;

    if (auto locked = wp.lock()) {
        std::cout << *locked << std::endl;
    } else {
        std::cout << "Object no longer exists.\n";
    }
}
\end{lstlisting}

\section*{Comparison Table}
\begin{tabular}{|l|l|l|l|}
\hline
Pointer Type & Ownership & Reference Counted & Auto Deletes \\
\hline
\texttt{unique\_ptr} & Exclusive & No & Yes \\
\texttt{shared\_ptr} & Shared & Yes & Yes (last one) \\
\texttt{weak\_ptr} & None & No & No \\
\hline
\end{tabular}

\section*{Summary}
\begin{itemize}
  \item Use \texttt{unique\_ptr} when only one owner exists.
  \item Use \texttt{shared\_ptr} when ownership is shared.
  \item Use \texttt{weak\_ptr} to observe an object managed by \texttt{shared\_ptr} without extending its lifetime.
\end{itemize}

\end{document}